---
title: "Team Final Project"
output: html_document
date: '2022-04-29'
---

1. Load the libraries that you will use in this project. 

```{r}
#install.packages('dplyr')
#install.packages('magrittr')
#install.packages('lubridate')
#install.packages('stringr')
#install.packages('tidyr')
library(dplyr)
library(magrittr)
library(lubridate)
library(stringr)
library(tidyr)
```

2. Read in the data from the ities.csv and max_temp.csv files as dataframe objects df and df_weather_weekly respectively.  

```{r}
df <- read.csv("ities.csv"
               , stringsAsFactors = T)
df_weather_weekly <- read.csv("max_temp.csv")
```

3. (1 point) Use an appropriate R function to display the number of rows and columns in the df and df_weather dataframes.

```{r}
dim(df)
dim(df_weather_weekly)
```

4. (4 points) Using any functions you want, make any necessary datatype conversions to columns in df and/or df_weather_weekly. Display the structures of the dataframes, df and df_weather_weekly to verify that the datatype conversions, if any, worked. Below the output, comment on why you did or did not make datatype conversions. (You may want to come back to this task and update it once you start working on the subsequent tasks.)  

```{r}
#Section 1
df$Department <- str_to_lower(df$Department)
df$LineItem <- str_to_lower(df$LineItem)
df$Category <- str_to_lower(df$Category)

#Section 2
df_weather_weekly <- df_weather_weekly %>%
  relocate(WeekStarting, .before = Monday)

#Section 3
df_weather_weekly$Week <- NULL

#Section 4
df$Date <- mdy(df$Date)
df_weather_weekly$WeekStarting <- mdy(df_weather_weekly$WeekStarting)

#Section 5
df <- df %>%
  arrange(Date)

```

*1. Section 1 : convert the Department, LineItem and Category to lower case to make it easier to utilize the data by conforming.*
*2. Section 2: Relocate the WeekStarting as the first column to join the data easier in the suceeeding steps.*
*3. Section 3: Delete Week column since it's repeating the row numbers to simplify the dataframe.*
*4. Section 4: Convert the factors or characters to date to tidy the data frame for ease of use in the succeeding steps.*
*5. Section 5: Arrange the data by acending order to make it similar to df_weather_weekly*

5. (1 point) Display summaries of the columns in the dataframes, df and df_weather_weekly. 

```{r}
summary(df)
summary(df_weather_weekly)
```

6. (3 points) Display the count of missing values in each column of df. Choose to remove or impute the missing values and defend that choice in 1-2 sentences for each column with missing values. 

```{r}
#Imputing the missing values to median price.
summary(df$Price, df$TotalDue)
df <- df %>%
  mutate(
    Price = ifelse(is.na(Price), median(Price, na.rm = T), Price)
    , TotalDue = Price * Quantity
  )
```

*All three variable Means are higher than medians. Which means the distribution is positively skewed. That's why missing values were imputed using median values to dampen the effect of the outliers.* 

7. (4 points) Let’s now explore whether there is a relationship in daily sales among the quantity purchased, the price paid, and the maximum temperature. To do so, we will want to aggregate the data at the daily level. Thus, please start by creating a new dataset df_daily by aggregating the dataframe df at the daily level. The dataframe df_daily must have the columns Quantity and TotalDue from df summarized at the daily level. Display a summary of df_daily. In one to two sentences defend the choice of summary measure (sum, mean, or other) used to aggregate the values in each column. 

```{r}
#Making df_daily from df.
df_daily <- df %>%
  mutate(
    date = round_date(Date, 'day')
  ) %>%
  group_by(Date) %>%
  summarise(
    AvgPrice = mean(Price, na.rm = T)
    , MedPrice = median(Price, na.rm = T)
    , Quantity = n()
    , TotalDue = sum(TotalDue, na.rm = T)
  ) %>%
  ungroup()
summary(df_daily)
```

*The df_daily shows required variables of daily total Quantity and TotalDue. It also includes mean and median price to show the observable patterns on daily basis such as $10.990 being the common median price for a lot of days.*

8. (4 points) Create a new dataframe, df_weather_daily, by pivoting the dataframe  df_weather_weekly from wide to long such that names of the seven columns (Monday:Sunday) are in a new column day and the values from those seven columns (Monday:Sunday) are in a new column max_temp. Display the first 10 rows of df_weather_daily to verify that the pivot worked.

```{r}
#Converting df_weather_weekly to daily by pivoting the weekly weather dataframe.
df_weather_daily <- df_weather_weekly %>%
  pivot_longer(cols = c(Monday:Sunday)
              , names_to = 'DayoftheWeek'
              , values_to = 'Max Temp')

#Proof the pivot worked.
df_weather_daily[1:10,]
```

9. (4 points) Update the df_weather_daily dataframe by adding a column, Date, that corresponds to the value in the day column. For example, if the WeekStarting value is “2016-01-03”, then the value in the Date column for the Monday that comes after “2016-01-03” should have a value of “2016-01-04”. Arrange df_weather_daily in ascending order of Date and display the first 10 rows to verify that it worked. Beneath the output comment on how you know that the creation of the Date column worked. 

```{r}
#Adding Date column from 2016-01-03 to 2016-12-31.
df_weather_daily$Date <- seq(as.Date('2016-01-03'), as.Date('2016-12-31'), 'days')

#Relocate Date column to the first column.
df_weather_daily <- df_weather_daily %>%
  relocate(Date, .before = WeekStarting)

#Show first 10 rows and 3 columns to verify that the creation of the Date column worked.
df_weather_daily[1:10,1:3]
```
*Referencing the first and last 10 lows of the Date column to WeekStarting and DayoftheWeek columbs, it has observed that the creation of the Date column worked.*


10. (2 points) Join the dataframes df_daily and df_weather_daily into one dataframe df_final such that only the rows that are in both df_daily and  df_weather_daily are in df_final. Arrange df_final in ascending order of Date. Then display something to show that the join worked. Briefly identify how that output helps you know that the join worked.   

```{r}
#Join the dataframes in a way that only the rows that are in both dataframes in acending order of date.
df_final <- df_daily %>%
  inner_join(df_weather_daily, by = 'Date')

#Disable WeekStarting on the df_final
df_final$WeekStarting <- NULL

#Relocate DayoftheWeek column to the second column.
df_final <- df_final %>%
  relocate(DayoftheWeek, .after = Date)

#Output Back-test
summary(df_final$Date)
df_daily[df_daily$Date >= '2016-01-05' & df_daily$Date <= '2016-12-31',]
```

*Looking at the summary output of df_final, it shows the min to max dates between 2016-01-05 to 2016-12-31. When df_daily observations are filtered between those dates, it shows 311 days. Which is the same as df_final. This proves that the join worked.*

11. (4 points) Display a pairplot like the one demonstrated towards the end of the module 2 video, “Getting to Know Your Data 3: Summary Statistics for Each Column, and Quick Plots”. Create this pairplot for three columns from df_final: Quantity, TotalDue, and max_temp. Below the output, identify whether there appears to be a two-way relationship for each of the two-way combination of variables: Quantity & TotalDue, Quantity & max_temp, and TotalDue & max_temp. 

```{r}
plot(df_final[,c('Quantity', 'TotalDue', 'Max Temp')], cex = 0.3)
```

### Two-Way Relationship Comparison
1. Quantity & TotalDue:
    *A positive correlation can be observed between these two variables.*
2. Quantity & Max_Temp:
    *A weak or neutral correlation can be observed between these two variables.* 
3. TotalDue & Max_Temp:
    *A weak or neutral correlation between these two variables.*

