---
title: "Divvy"
author: "Michael R. Mercado"
date: "6/6/2022"
output:
  html_document: default
  pdf_document: default
---

## LOAD PACKAGES
```{r}
library(lubridate)
library(stringi)
library(magrittr)
library(tidyverse) # For lots of data wrangling tasks
```

## Read in the data.
```{r}
df_01_22 <- read.csv('Jan22_divvy_tripdata.csv', stringsAsFactors = F, header = T)    
# Reading in the January 2022 Divvy data as a dataframe called df_01_22
df_02_22 <- read.csv('Feb22_divvy_tripdata.csv', stringsAsFactors = F, header = T)     
# Reading in the February 2022 Divvy data as a dataframe called df_02_22
df_03_22 <- read.csv('Mar22_divvy_tripdata.csv', stringsAsFactors = F, header = T)     
# Reading in the March 2022 Divvy data as a dataframe called df_03_22
df_04_22 <- read.csv('Apr22_divvy_tripdata.csv', stringsAsFactors = F, header = T)     
# Reading in the April 2022 Divvy data as a dataframe called df_04_22
df_05_22 <- read.csv('May22_divvy_tripdata.csv', stringsAsFactors = F, header = T)     
# Reading in the May 2022 Divvy data as a dataframe called df_05_22
```

## Merge the Data
```{r}
df_Merged <- rbind(df_01_22, df_02_22, df_03_22, df_04_22, df_05_22)
```

## Structure of the Data
```{r}
str(df_Merged) 
```

## Check Missing Valaues
```{r}
colSums(is.na(df_Merged)) # Check if there are missing values 
```

## Remove Missing Valaues
```{r}
df_cleaned <- na.omit(df_Merged)     # Removes rows containing missing values
colSums(is.na(df_cleaned)) # Check to make sure the missing values are no more.
```
## Convert the Started_At column from character string object into a date-time object
```{r}
df_cleaned <- df_cleaned %>% # Object will be converted from a character string to a date time object
    mutate(
    started_at = ymd_hms(started_at)
   ) 
```

## Convert the Ended_At column from character string object into a date-time object
```{r}
df_cleaned <- df_cleaned %>% # Object will be converted from a character string to a date time object
    mutate(
    ended_at = ymd_hms(ended_at)
   ) 
```

## Unique Valaues for Rideable Type
```{r}
unique(df_cleaned$rideable_type)
```

## Unique Valaues for Member Casual
```{r}
unique(df_cleaned$member_casual)
```

## Create Two Columns for Calculated Trip Duration, One Column in Trip Minutes and Second Column in Trip Hours
```{r}
df_triptime <- df_cleaned

df_triptime <- df_triptime %>%
  mutate(
          trip_time_min = as.numeric(difftime(df_triptime$ended_at, df_triptime$started_at), units = "mins"),
          trip_time_hrs = as.numeric(difftime(df_triptime$ended_at, df_triptime$started_at), units = "hours")
         )
```

## For Any Trip With a Duration Less Than or Equal To Zero Minutes or Greater Than 3 Hours Replace with NA's
```{r}
df_triptime$trip_time_hrs <- replace(df_triptime$trip_time_hrs, which(df_triptime$trip_time_hrs <= 0.167), NA)
df_triptime$trip_time_hrs <- replace(df_triptime$trip_time_hrs, which(df_triptime$trip_time_hrs > 3), NA)
```

## Remove Rows with Missing Valaues
```{r}
df_triptime <- na.omit(df_triptime)     # Removes rows containing missing values
colSums(is.na(df_triptime)) # Check to make sure the missing values are no more.
```

## Structure of the Data
```{r}
str(df_triptime) 
```
## Box plot For Trip Time By User Type For Trips <= 3 hours
```{r}
ggplot(df_triptime, aes(x=member_casual, y=trip_time_hrs)) + 
  geom_boxplot() +
      labs(title = 'Box Plot-Trip Time Duration (Max 3 Hours) By Type of User')
```
## Filter Data for Trips Up to 1 Hour
```{r}
df_onehourtrips <- df_triptime
df_onehourtrips$trip_time_hrs <- replace(df_onehourtrips$trip_time_hrs, which(df_onehourtrips$trip_time_hrs <= 0.167), NA)
df_onehourtrips$trip_time_hrs <- replace(df_onehourtrips$trip_time_hrs, which(df_onehourtrips$trip_time_hrs > 1), NA)
```

## Remove Rows with Missing Valaues
```{r}
df_onehourtrips <- na.omit(df_onehourtrips)     # Removes rows containing missing values
colSums(is.na(df_onehourtrips)) # Check to make sure the missing values are no more.
```

## Box plot For Trip Time By User Type For Trips <= 1 hour
```{r}
ggplot(df_onehourtrips, aes(x=member_casual, y=trip_time_hrs)) + 
  geom_boxplot() +
      labs(title = 'Box Plot-Trip Time Duration (Max 1 Hour) By Type of User')
```

## Filter Data for Only Casual Riders
```{r}
df_casual <- df_triptime %>% 
  filter(df_triptime$member_casual=="casual")
```

## Export Output of Causal Rider Data
```{r}
write.csv(df_casual,"CasualCleaned.csv", row.names = TRUE)
```

## Filter Data for Only Member Riders
```{r}
df_member <- df_triptime %>% 
  filter(df_triptime$member_casual=="member")
```

## Group Data to See Distribution for Stations Among Casual Riders from their Start Stations
```{r}
df_casual_startedat_stations <- df_casual %>% group_by(start_station_id) %>% count(start_station_id) %>% arrange(desc(n)) %>% slice(1:10) 
```

## Group Data to See Distribution for Stations Among Casual Riders Form The End Stations
```{r}
df_casual_endedat_stations <- df_casual %>% group_by(end_station_id) %>% count(end_station_id) %>% arrange(desc(n)) %>% slice(1:10) 
```

## Export Data Set 1
```{r}
write.csv(df_casual_startedat_stations,"Casual_Start_Stations.csv", row.names = TRUE)
```

## Export Data Set 2
```{r}
write.csv(df_casual_endedat_stations,"Casual_End_Stations.csv", row.names = TRUE)
```

## Box plot For Trip Time By User Type
```{r}
ggplot(df_member, aes(x=rideable_type, y=trip_time_hrs)) + 
  geom_boxplot() +
      labs(title = 'Box Plot-Trip Time Duration By Type of Bike Used for Members')
```

## Box plot For Trip Time By Bike Type
```{r}
ggplot(df_casual, aes(x=rideable_type, y=trip_time_hrs)) + 
  geom_boxplot() +
      labs(title = 'Box Plot-Trip Time Duration By Type of Bike Used for Casual Riders')
```


